---
title: "Laboratory Session - April 28, 2022"
output: html_notebook
editor_options:
  chunk_output_type: inline
---

# Paolo Zinesi 2053062

```{r}
set.seed(12345)

library(tidyverse)
library(gridExtra)

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

## Exercise 1

The number of particles emitted by a radioactive source during a fixed interval of time (∆t = 10 s) follows a Poisson distribution on the parameter $\mu$. The number of particles observed during consecutive time intervals is:
$$\{ 4, 1, 3, 1, 3 \}$$

We use priors that are conjugate to the likelihood. For a Poisson process the conjugate prior is a Gamma distribution, which PDF is $\text{Gamma} \left(y\,|\,\alpha,\lambda \right) = \dfrac{\lambda^{\alpha}}{\Gamma{(\alpha)}}\, y^{\alpha-1}\, e^{-\lambda\, y}$. If the prior has the form $\text{Gamma} \left(\alpha,\lambda \right)$ and we have $n$ observations with results $\{y_i\}$, the posterior has the form $\text{Gamma}\left(\alpha + \sum_i y_i, \,\lambda + n \right)$. The mean of a $\text{Gamma} \left(\alpha,\lambda \right)$ distribution is $\dfrac{\alpha}{\lambda}$ and the variance is $\dfrac{\alpha}{\lambda^2}$


```{r}
res <- c(4, 1, 3, 1, 3)
mu <- seq(0, 10, length.out=1000)
N_dl <- length(mu) - 1
dl <- (max(mu) - min(mu)) / N_dl
```

### Exercise 1a

Suppose a uniform prior distribution for the parameter $\mu$.
Determine and draw the posterior distribution for $\mu$, given the data $D=\{y_i\}$.

A uniform prior has the form $\text{Gamma} \left(1,0 \right)$, so the posterior has the form $\text{Gamma} \left(1+\sum_i y_i, \,0 + n \right)$.

```{r}
# prior
log_prior <- ifelse(mu>=0, 0, -Inf)

# likelihood
log_likl <- 0
for (r_ in res) {
  log_likl <- log_likl + dpois(x=r_, lambda=mu, log=TRUE)
}

# posterior
post_U <- exp(log_likl + log_prior)
post_U <- post_U / (dl * sum(post_U))

# plot posterior
post_U_plt <- ggplot() +
              labs(title=paste0("Posterior = Gamma(", 1+sum(res),",", length(res),")", collapse=""), subtitle="Corresponding to a uniform prior") +
              geom_line(data = data.frame(mu=mu, post=post_U), aes(x=mu, y=post), color="steelblue", size=1) +
              scale_y_continuous(name = expression(paste("P( ", mu, " | D,M )"))) +
              scale_x_continuous(name = expression(mu)) +
              theme_bw() +
              theme(plot.title=element_text(hjust=0.5), plot.subtitle=element_text(hjust=0.5))
post_U_plt
```



Evaluate mean, median and variance, both analytically and numerically in R.
```{r}
mean_num_U <- sum(mu*post_U) * dl
median_num_U <- mu[sum(cumsum(post_U * dl) <= 0.5) + 1]
var_num_U <- sum(mu * mu * post_U * dl) - (mean_num_U * mean_num_U)
mean_an_U <- (1 + sum(res)) / length(res)
median_an_U <- qgamma(0.5, shape=1+sum(res), rate=length(res))
var_an_U <- (1 + sum(res)) / length(res)**2

cat("Mean numerical = ", mean_num_U, "; ", "Mean analytical = ", mean_an_U, "\n", sep="")
cat("Median numerical = ", median_num_U, "; ", "Median analytical = ", median_an_U, "\n", sep="")
cat("Variance numerical = ", var_num_U, "; ", "Variance analytical = ", var_an_U, "\n", sep="")
```

```{r}
post_U_plt <- post_U_plt +
              geom_vline(mapping=aes(xintercept = mean_an_U, color="Mean"), linetype="dashed", size=0.5) +
              geom_vline(mapping=aes(xintercept = median_an_U, color="Median"), linetype="dashed", size=0.5) +
              scale_color_manual(values = cbPalette, name="Values")
post_U_plt
```



### Exercise 1b

Suppose a Jeffrey’s prior for the parameter $\mu$.
Determine and draw the posterior distribution for $\mu$, given the data.

A Jeffrey's prior has the form $\text{Gamma} \left(\dfrac{1}{2},0 \right)$, so the posterior has the form $\text{Gamma} \left(\dfrac{1}{2}+\sum_i y_i, \,0 + n \right)$.

```{r}
# prior
log_prior <- ifelse(mu > 0, -0.5*log(mu), -Inf)

# likelihood
log_likl <- 0
for (r_ in res) {
  log_likl <- log_likl + dpois(x=r_, lambda=mu, log=TRUE)
}

# posterior
post_J <- exp(log_likl + log_prior)
post_J <- post_J / (dl * sum(post_J))

# plot posterior
post_J_plt <- ggplot() +
              labs(title=paste0("Posterior = Gamma(", 0.5+sum(res),",", length(res),")", collapse=""), subtitle="Corresponding to a Jeffrey's prior") +
              geom_line(data = data.frame(mu=mu, post=post_J), aes(x=mu, y=post), color="steelblue", size=1) +
              scale_y_continuous(name = expression(paste("P( ", mu, " | D,M )"))) +
              scale_x_continuous(name = expression(mu)) +
              theme_bw() +
              theme(plot.title=element_text(hjust=0.5), plot.subtitle=element_text(hjust=0.5))
post_J_plt
```


Evaluate mean, median and variance, both analytically and numerically in R.
```{r}
mean_num_J <- sum(mu*post_J) * dl
median_num_J <- mu[sum(cumsum(post_J * dl) <= 0.5) + 1]
var_num_J <- sum(mu * mu * post_J * dl) - (mean_num_J * mean_num_J)
mean_an_J <- (0.5 + sum(res)) / length(res)
median_an_J <- qgamma(0.5, shape=0.5+sum(res), rate=length(res))
var_an_J <- (0.5 + sum(res)) / length(res)**2

cat("Mean numerical = ", mean_num_J, "; ", "Mean analytical = ", mean_an_J, "\n", sep="")
cat("Median numerical = ", median_num_J, "; ", "Median analytical = ", median_an_J, "\n", sep="")
cat("Variance numerical = ", var_num_J, "; ", "Variance analytical = ", var_an_J, "\n", sep="")
```


```{r}
post_J_plt <- post_J_plt +
              geom_vline(mapping=aes(xintercept = mean_an_J, color="Mean"), linetype="dashed", size=0.5) +
              geom_vline(mapping=aes(xintercept = median_an_J, color="Median"), linetype="dashed", size=0.5) +
              scale_color_manual(values = cbPalette, name="Values")
post_J_plt
```

### Exercise 1c
Evaluate a 95% credibility interval for the results obtained with both priors. Compare the result with that obtained using a normal approximation for the posterior distribution, with the same mean and standard deviation.

```{r}
# credibility interval for both posteriors
CI_U <- c(qgamma(0.025, shape=1+sum(res), rate=length(res)), qgamma(0.975, shape=1+sum(res), rate=length(res)))
CI_U_gauss <- c(qnorm(0.025, mean=mean_an_U, sd=sqrt(var_an_U)), qnorm(0.975, mean=mean_an_U, sd=sqrt(var_an_U)))
CI_J <- c(qgamma(0.025, shape=0.5+sum(res), rate=length(res)), qgamma(0.975, shape=0.5+sum(res), rate=length(res)))
CI_J_gauss <- c(qnorm(0.025, mean=mean_an_J, sd=sqrt(var_an_J)), qnorm(0.975, mean=mean_an_J, sd=sqrt(var_an_J)))

cat("Credibility interval corresponding to a uniform prior:", CI_U, "\n")
cat("Credibility interval corresponding to a uniform prior with a normal approximation:", CI_U_gauss, "\n")
cat("Credibility interval corresponding to a Jeffrey's prior:", CI_J, "\n")
cat("Credibility interval corresponding to a Jeffrey's prior with a normal approximation:", CI_J_gauss, "\n")
```


```{r, warning=FALSE, fig.width=5, fig.height=3}
post_U_plt <- post_U_plt + geom_area(data=data.frame(mu=mu, post=post_U),
                                     aes(x=ifelse(mu>=CI_U[1] & mu<=CI_U[2], mu, NA), y=post, fill="Uniform"),
                                     alpha = 0.2) +
                           scale_fill_manual(values="steelblue", name="95% CI")
post_J_plt <- post_J_plt + geom_area(data=data.frame(mu=mu, post=post_J),
                                     aes(x=ifelse(mu>=CI_J[1] & mu<=CI_J[2], mu, NA), y=post, fill="Jeffrey"),
                                     alpha = 0.2) +
                           scale_fill_manual(values="steelblue", name="95% CI")

grid.arrange(post_U_plt, post_J_plt, nrow=2)
```


## Exercise 2

Given the problem of the lightouse discussed last week, study the case in which both the position along the shore ($\alpha$) and the distance out at sea ($\beta$) are unknown.

We generate random $\theta \in \left[ -\dfrac{\pi}{2}, \dfrac{\pi}{2} \right]$ and we calculate the corresponding $x$ following the relation
$$x_k = \alpha + \beta\, \tan{\theta_k}.$$

The prior considered is constant in the rectangle $(\alpha,\beta) \in \left[ -2\, \text{km}, 2\, \text{km} \right] \times \left[ 0\, \text{km}, 5\, \text{km} \right]$ and null elsewhere. The posterior of the problem is thus equivalent to the likelihood in this range and null elsewhere. The log-likelihood used for this problem is 
$$\log{\mathcal{L}(D\, |\, \alpha,\beta,M)} = \sum_i \log \left[ \dfrac{\beta}{\beta^2 + (x_i - \alpha)^2} \right]$$ 
where $D=\{x_i\}$ is the set of x-values measured.

```{r}
# fix true alpha and beta
alpha_true <- 1 #km
beta_true <- 1 #km
alpha_min <- -2 #km
alpha_max <- 2 #km
beta_max <- 5 #km

# define grid
alpha0 <- seq(alpha_min, alpha_max, length.out=1001)
d_alpha <- (alpha_max - alpha_min) / 1000
beta0 <- seq(0, beta_max, length.out=1001)
d_beta <- (beta_max - 0) / 1000
grid0 <- expand_grid(alpha=alpha0, beta=beta0)

# draw samples
n_max <- 1000
lighthouse_samples <- alpha_true + beta_true * tan(runif(n_max, min=-pi/2, max=pi/2))
```


```{r}
# function that computes posteriors for the lighthouse problem
compute_posterior_lighthouse <- function(grid_, da_, db_, samples_, marginalize="") {

  # log-posterior
  post_df <- grid_ %>% mutate(log_post = 0)
  for (x_ in samples_) {
    post_df <- post_df %>%
               mutate(log_post = log_post + log(beta/(beta**2 + (x_ - alpha)**2)))
  }
  post_df <- post_df %>% mutate(log_post = log_post - max(log_post))

  # unnormalized posterior
  post_df$post <- exp(post_df$log_post)


  if(marginalize=="alpha") {

    # marginalize over beta and normalize the marginalized posterior
    post_alpha <- post_df %>% group_by(alpha) %>% summarise(post = sum(post, na.rm = TRUE))
    post_alpha$post <- post_alpha$post / (sum(post_alpha$post) * da_)

    return(post_alpha)
  }
  else if (marginalize=="beta") {

    # marginalize over alpha and normalize the marginalized posterior
    post_beta <- post_df %>% group_by(beta) %>% summarise(post = sum(post, na.rm = TRUE))
    post_beta$post <- post_beta$post / (sum(post_beta$post) * db_)

    return(post_beta)
  }
  else {

    # no marginalization
    return(post_df)
  }

}
```



```{r, fig.height=5, fig.width=6}
plt_base <- ggplot() +
            geom_vline(mapping=aes(xintercept = alpha_true), colour="firebrick", linetype="dashed", size=1) +
            scale_x_continuous(name="Alpha") +
            scale_y_continuous(name=expression(paste("P( ", alpha, " | D,M )"))) +
            theme_bw() +
            theme(plot.title=element_text(hjust=0.5))


# plot alpha posterior
grid.arrange(
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1], marginalize = "alpha"),
                       aes(x=alpha, y=post)) +
             labs(title="Marginalized posterior with 1 sample"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:2], marginalize = "alpha"),
                       aes(x=alpha, y=post)) +
             labs(title="Marginalized posterior with 2 samples"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:5], marginalize = "alpha"),
                       aes(x=alpha, y=post)) +
             labs(title="Marginalized posterior with 5 samples"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:10], marginalize = "alpha"),
                       aes(x=alpha, y=post)) +
             labs(title="Marginalized posterior with 10 samples"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:20], marginalize = "alpha"),
                       aes(x=alpha, y=post)) +
             labs(title="Marginalized posterior with 20 samples"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:50], marginalize = "alpha"),
                       aes(x=alpha, y=post)) +
             labs(title="Marginalized posterior with 50 samples"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:100], marginalize = "alpha"),
                       aes(x=alpha, y=post)) +
             labs(title="Marginalized posterior with 100 samples"),
  nrow=4)

```

```{r, fig.height=5, fig.width=6}
plt_base <- ggplot() +
            geom_vline(mapping=aes(xintercept = beta_true), colour="firebrick", linetype="dashed", size=1) +
            scale_x_continuous(name="Beta") +
            scale_y_continuous(name=expression(paste("P( ", beta, " | D,M )"))) +
            theme_bw() +
            theme(plot.title=element_text(hjust=0.5))


# plot beta posterior
grid.arrange(
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1], marginalize = "beta"),
                       aes(x=beta, y=post)) +
             labs(title="Marginalized posterior with 1 sample"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:2], marginalize = "beta"),
                       aes(x=beta, y=post)) +
             labs(title="Marginalized posterior with 2 samples"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:5], marginalize = "beta"),
                       aes(x=beta, y=post)) +
             labs(title="Marginalized posterior with 5 samples"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:10], marginalize = "beta"),
                       aes(x=beta, y=post)) +
             labs(title="Marginalized posterior with 10 samples"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:20], marginalize = "beta"),
                       aes(x=beta, y=post)) +
             labs(title="Marginalized posterior with 20 samples"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:50], marginalize = "beta"),
                       aes(x=beta, y=post)) +
             labs(title="Marginalized posterior with 50 samples"),
  plt_base + geom_line(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:100], marginalize = "beta"),
                       aes(x=beta, y=post)) +
             labs(title="Marginalized posterior with 100 samples"),
  nrow=4)
```




```{r, fig.height=7, fig.width=5}
plt_base <- ggplot() +
            scale_x_continuous(name="Alpha") +
            scale_y_continuous(name="Beta") +
            theme(plot.title=element_text(hjust=0.5)) +
            theme_bw()

grid.arrange(
  plt_base + geom_contour_filled(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:2]),
                    mapping = aes(x=alpha, y=beta, z=post), show.legend=FALSE) +
             geom_vline(mapping=aes(xintercept = alpha_true), colour="firebrick", linetype="dashed", size=1) +
             geom_hline(mapping=aes(yintercept = beta_true), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Posterior with 2 samples") +
             theme(plot.title=element_text(hjust=0.5)),
  plt_base + geom_contour_filled(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:5]),
                    mapping = aes(x=alpha, y=beta, z=post), show.legend=FALSE) +
             geom_vline(mapping=aes(xintercept = alpha_true), colour="firebrick", linetype="dashed", size=1) +
             geom_hline(mapping=aes(yintercept = beta_true), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Posterior with 5 samples") +
             theme(plot.title=element_text(hjust=0.5)),
  plt_base + geom_contour_filled(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:10]),
                    mapping = aes(x=alpha, y=beta, z=post), show.legend=FALSE) +
             geom_vline(mapping=aes(xintercept = alpha_true), colour="firebrick", linetype="dashed", size=1) +
             geom_hline(mapping=aes(yintercept = beta_true), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Posterior with 10 samples") +
             theme(plot.title=element_text(hjust=0.5)),
  plt_base + geom_contour_filled(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:20]),
                    mapping = aes(x=alpha, y=beta, z=post), show.legend=FALSE) +
             geom_vline(mapping=aes(xintercept = alpha_true), colour="firebrick", linetype="dashed", size=1) +
             geom_hline(mapping=aes(yintercept = beta_true), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Posterior with 20 samples") +
             theme(plot.title=element_text(hjust=0.5)),
  plt_base + geom_contour_filled(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:50]),
                    mapping = aes(x=alpha, y=beta, z=post), show.legend=FALSE) +
             geom_vline(mapping=aes(xintercept = alpha_true), colour="firebrick", linetype="dashed", size=1) +
             geom_hline(mapping=aes(yintercept = beta_true), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Posterior with 50 samples") +
             theme(plot.title=element_text(hjust=0.5)),
  plt_base + geom_contour_filled(data=compute_posterior_lighthouse(grid0, d_alpha, d_beta, lighthouse_samples[1:100]),
                    mapping = aes(x=alpha, y=beta, z=post), show.legend=FALSE) +
             geom_vline(mapping=aes(xintercept = alpha_true), colour="firebrick", linetype="dashed", size=1) +
             geom_hline(mapping=aes(yintercept = beta_true), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Posterior with 100 samples") +
             theme(plot.title=element_text(hjust=0.5)),
  nrow=3)
```

```{r}
# find most probables (alpha,beta) for each iteration
max_prob_alpha <- c()
max_prob_beta <- c()
joint_post_df <- grid0 %>% mutate(log_post = 0)

for(x_ in lighthouse_samples[1:100]) {
  joint_post_df <- joint_post_df %>%
                   mutate(log_post = log_post + log(beta/(beta**2 + (x_ - alpha)**2)))
  
  max_indx <- which.max(joint_post_df$log_post)
  max_prob_alpha <- c(max_prob_alpha, joint_post_df[[max_indx, "alpha"]])
  max_prob_beta <- c(max_prob_beta, joint_post_df[[max_indx, "beta"]])
}

# plot succession of most probable values
plt_max_prob_evol <- ggplot(data=data.frame(Alpha=max_prob_alpha, Beta=max_prob_beta),
                            mapping=aes(x=Alpha, y=Beta)) +
                     geom_path() +
                     geom_vline(mapping=aes(xintercept = alpha_true), colour="firebrick", linetype="dashed", size=1) +
                     geom_hline(mapping=aes(yintercept = beta_true), colour="firebrick", linetype="dashed", size=1) +
                     geom_point() +
                     geom_point(aes(x=alpha_true, y=beta_true), colour="firebrick") +
                     geom_text(data=data.frame(alpha=max_prob_alpha[1:10], beta=max_prob_beta[1:10], labels=1:10),
                               mapping=aes(x=alpha+0.05, y=beta+0.05, label = labels)) +
                     labs(title=expression(paste("Evolution of the most probables (", alpha, " , ", beta, ") as a function of number of data samples"))) +
                     theme_bw() +
                     theme(plot.title=element_text(hjust=0.5))
                     

plt_max_prob_evol
```

**As expected, by updating iteratively the posterior with the data, the posterior distribution becomes more and more peaked around the true value of the parameters $\alpha, \beta$.**



## Exercise 3

Given the Signal over Background example discussed last week, analyze and discuss the following cases.

The signal is modeled by the following function:

$$ S\,(x\,|\,A,B,x_0,w,\Delta t) = \Delta t \left( A\, e^{-\dfrac{1}{2} \left(\dfrac{x-x_0}{w}\right)^2 } +B \right).$$

The prior considered is constant in the rectangle $(A,B) \in \left[ 0, 5 \right] \times \left[ 0.5, 1.5 \right]$ and null elsewhere. The posterior of the problem is thus equivalent to the likelihood in this range and null elsewhere. The log-likelihood used for this problem is 
$$\log{\mathcal{L}(D\, |\, A,B,M)} = \sum_i \left[ N_i \ln{S\,(x_i\,|\,A,B)} - S\,(x_i\,|\,A,B) - \ln{(N_i!)} \right]$$
where $D=\{N_i\}$ is the set of counts measured in the positions $\{x_i\}$.
```{r}
# (starting) model definitions
x0 <- 0
w <- 1
A_true <- 2
B_true <- 1
dt <- 5
x_grid <- seq(-7*w, 7*w, by=0.5*w)

# define grid for parameters A,B
A_max <- 5
B_min <- 0.5
B_max <- 1.5
A0 <- seq(0, A_max, length.out=1001)
dA <- (A_max - 0) / 1000
B0 <- seq(B_min, B_max, length.out=1001)
dB <- (B_max - B_min) / 1000
grid0 <- expand_grid(A=A0, B=B0)


# signal definition
signal <- function(x, A, B, x0, w, dt) {
  dt * (A * exp(-0.5*((x-x0)/w)**2) + B)
}

# true signal
signal_true_df <- tibble(x=x_grid, Signal=signal(x_grid, A_true, B_true, x0, w, dt))
plt_SB <- ggplot() +
          geom_line(data=signal_true_df, mapping=aes(x=x, y=Signal)) +
          labs(title="Real signal vs measured signal with A/B=2 and w=1") +
          scale_y_continuous(name = expression(paste("Signal (x | A, B, ", x[0], ", w, ", Delta, "t, M)"))) +
          theme_bw() +
          theme(plot.title=element_text(hjust=0.5))
          

# sample signals
signal_samples <- rpois(length(x_grid), signal_true_df$Signal)
signal_samples_df <- tibble(x=x_grid, N=signal_samples)
plt_SB <- plt_SB +
          geom_col(data=signal_samples_df, mapping=aes(x=x, y=N), alpha=0.3)

plt_SB
```



```{r}
# function that computes the posterior for the Signal/Background problem,
# it computes the posterior for different values of the parameters
# w, A_true, B_true, x0, dt
compute_posterior_SB <- function(x0=0, w=1, A_true=2, B_true=1, dt=5,
                                 A_max=5, B_min=0.5, B_max=1.5,
                                 marginalize="") {
  
  # reproducibility
  set.seed(12345)
  
  # define grid
  grid_x_ <- seq(-7*w, 7*w, by=0.5*w)
  grid_ <- expand_grid(A=seq(0, A_max, length.out=1001), B=seq(B_min, B_max, length.out=1001))
  dA_ <- (A_max - 0) / 1000
  dB_ <- (B_max - B_min) / 1000
  
  
  # true signal
  signal_true <- signal(grid_x_, A_true, B_true, x0, w, dt)
  
  # sample signals
  samples_ <- rpois(n=length(grid_x_), lambda=signal_true)
  
  
  
  # log-posterior
  post_df <- grid_ %>% mutate(log_post = 0)
  for (i in 1:length(samples_)) {
    post_df <- post_df %>%
               mutate(log_post = log_post + dpois(x=samples_[i], lambda = signal(grid_x_[i], A, B, x0=x0, w=w, dt=dt), log=TRUE))
  }
  post_df <- post_df %>% mutate(log_post = log_post - max(log_post))

  # unnormalized posterior
  post_df$Posterior <- exp(post_df$log_post)
  
  
  
  if(marginalize=="A") {

    # marginalize over B and normalize the marginalized posterior
    post_A <- post_df %>% group_by(A) %>% summarise(Posterior = sum(Posterior, na.rm = TRUE))
    post_A$Posterior <- post_A$Posterior / (sum(post_A$Posterior) * dA_)

    return(post_A)
  }
  else if (marginalize=="B") {

    # marginalize over A and normalize the marginalized posterior
    post_B <- post_df %>% group_by(B) %>% summarise(Posterior = sum(Posterior, na.rm = TRUE))
    post_B$Posterior <- post_B$Posterior / (sum(post_B$Posterior) * dB_)

    return(post_B)
  }
  else {

    # no marginalization
    return(post_df)
  }
  
}
```


### Exercise 3a
Vary the sampling resolution used to generate the data, keeping the same sampling range.
Change the resolution w = {0.1, 0.25, 1, 2, 3} and check the effect on the results.


```{r}
plot_signal <- function(x0=0, w=1, A_true=2, B_true=1, dt=5,
                        A_max=5, B_min=0.5, B_max=1.5, xlim=c(-8*w,8*w)) {
  
  # reproducibility
  #set.seed(12345)
  
  # define grid
  grid_x_ <- seq(-7*w, 7*w, by=0.5*w)
  
  # true signal
  signal_true <- signal(grid_x_, A_true, B_true, x0, w, dt)
  signal_true_df <- tibble(x=grid_x_, Signal=signal_true)
  
  # sample signals
  signal_samples <- rpois(length(grid_x_), signal_true_df$Signal)
  signal_samples_df <- tibble(x=grid_x_, N=signal_samples)
  
  
  
  # plot
  plt <- ggplot() +
         geom_line(data=signal_true_df, mapping=aes(x=x, y=Signal)) +
         geom_col(data=signal_samples_df, mapping=aes(x=x, y=N), alpha=0.3) +
         labs(title=paste0("Real signal vs measured signal with A/B=", A_true/B_true, " and w=", w, collapse = "")) +
         scale_x_continuous(limits=xlim) +
         scale_y_continuous(name = expression(paste("Signal (x | A, B, ", x[0], ", w, ", Delta, "t, M)"))) +
         theme_bw() +
         theme(plot.title=element_text(hjust=0.5))
  
  return(plt)
}
```




```{r, fig.height=5, fig.width=6}
# plot real signals vs measured signal as a function of w
grid.arrange(
plot_signal(w=0.1, A_true=2, B_true=1, xlim=c(-22,22)),
plot_signal(w=0.25, A_true=2, B_true=1, xlim=c(-22,22)),
plot_signal(w=1, A_true=2, B_true=1, xlim=c(-22,22)),
plot_signal(w=2, A_true=2, B_true=1, xlim=c(-22,22)),
plot_signal(w=3, A_true=2, B_true=1, xlim=c(-22,22)),
nrow=3)
```


**As it is possible to see from the plots above, the change of the scale parameter $w$ if the sampling range remains fixed affects the spatial width of the signal but not the parameters $A,B$. The plots below show the independence of $A,B$ on $w$ when the sampling range is fixed. This effect occurs because the signal depends only on the ratio $x/w$ ($x_0=0$) and not singularly on the two variables. Since the x-grid is proportional to $w$, the signal generated are exactly the same. To see some changes in the plots we could fix the spatial (x-)width of the signal and vary the parameter $w$ in order to vary the number of channels that can collect the signal.**


```{r, fig.width=5}
# plot two posteriors as an example of invariance w.r.t. w parameter
grid.arrange(
  ggplot() + geom_contour_filled(data=compute_posterior_SB(w=0.1), mapping = aes(x=A, y=B, z=Posterior), show.legend=FALSE) +
             labs(title=paste("Posterior with A/B=2 and w=0.1")) +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  ggplot() + geom_contour_filled(data=compute_posterior_SB(w=3), mapping = aes(x=A, y=B, z=Posterior), show.legend=FALSE) +
             labs(title=paste("Posterior with A/B=2 and w=3")) +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  nrow=1)
```



### Exercise 3b
Change the ratio A/B used to simulate the data (keeping both positive in accordance with the prior). Check the effect on the results.

We investigate the ratios $A/B = \{ 0.25, 0.5, 1, 2, 4\}$ by fixing $B=1$ and choosing $A= \{ 0.25, 0.5, 1, 2, 4\}$.

```{r, fig.height=5, fig.width=6}
# plot real signals vs measured signal as a function of A/B
grid.arrange(
plot_signal(w=1, A_true=0.25, B_true=1),
plot_signal(w=1, A_true=0.5, B_true=1),
plot_signal(w=1, A_true=1, B_true=1),
plot_signal(w=1, A_true=2, B_true=1),
plot_signal(w=1, A_true=4, B_true=1),
nrow=3)
```



```{r, fig.height=5, fig.width=6}
# plot A posterior
grid.arrange(
  ggplot() + geom_line(data=compute_posterior_SB(A_true=0.25, B_true=1, marginalize = "A"),
                       aes(x=A, y=Posterior)) +
             geom_vline(mapping=aes(xintercept = 0.25), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Marginalized posterior with A/B=0.25 and w=1") +
             scale_y_continuous(name="P( A | D,M )") +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  ggplot() + geom_line(data=compute_posterior_SB(A_true=0.5, B_true=1, marginalize = "A"),
                       aes(x=A, y=Posterior)) +
             geom_vline(mapping=aes(xintercept = 0.5), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Marginalized posterior with A/B=0.5 and w=1") +
             scale_y_continuous(name="P( A | D,M )") +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  ggplot() + geom_line(data=compute_posterior_SB(A_true=1, B_true=1, marginalize = "A"),
                       aes(x=A, y=Posterior)) +
             geom_vline(mapping=aes(xintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Marginalized posterior with A/B=1 and w=1") +
             scale_y_continuous(name="P( A | D,M )") +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  ggplot() + geom_line(data=compute_posterior_SB(A_true=2, B_true=1, marginalize = "A"),
                       aes(x=A, y=Posterior)) +
             geom_vline(mapping=aes(xintercept = 2), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Marginalized posterior with A/B=2 and w=1") +
             scale_y_continuous(name="P( A | D,M )") +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  ggplot() + geom_line(data=compute_posterior_SB(A_true=4, B_true=1, marginalize = "A"),
                       aes(x=A, y=Posterior)) +
             geom_vline(mapping=aes(xintercept = 4), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Marginalized posterior with A/B=4 and w=1") +
             scale_y_continuous(name="P( A | D,M )") +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  nrow=3)
```



```{r, fig.height=5, fig.width=6}
# plot B posterior
grid.arrange(
  ggplot() + geom_line(data=compute_posterior_SB(A_true=0.25, B_true=1, marginalize = "B"),
                       aes(x=B, y=Posterior)) +
             geom_vline(mapping=aes(xintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Marginalized posterior with A/B=0.25 and w=1") +
             scale_y_continuous(name="P( B | D,M )") +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  ggplot() + geom_line(data=compute_posterior_SB(A_true=0.5, B_true=1, marginalize = "B"),
                       aes(x=B, y=Posterior)) +
             geom_vline(mapping=aes(xintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Marginalized posterior with A/B=0.5 and w=1") +
             scale_y_continuous(name="P( B | D,M )") +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  ggplot() + geom_line(data=compute_posterior_SB(A_true=1, B_true=1, marginalize = "B"),
                       aes(x=B, y=Posterior)) +
             geom_vline(mapping=aes(xintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Marginalized posterior with A/B=1 and w=1") +
             scale_y_continuous(name="P( B | D,M )") +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  ggplot() + geom_line(data=compute_posterior_SB(A_true=2, B_true=1, marginalize = "B"),
                       aes(x=B, y=Posterior)) +
             geom_vline(mapping=aes(xintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Marginalized posterior with A/B=2 and w=1") +
             scale_y_continuous(name="P( B | D,M )") +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  ggplot() + geom_line(data=compute_posterior_SB(A_true=4, B_true=1, marginalize = "B"),
                       aes(x=B, y=Posterior)) +
             geom_vline(mapping=aes(xintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Marginalized posterior with A/B=4 and w=1") +
             scale_y_continuous(name="P( B | D,M )") +
             theme_bw() +
             theme(plot.title=element_text(hjust=0.5)),
  nrow=3)
```


```{r, fig.height=7, fig.width=5}
plt_base <- ggplot() +
            theme_bw() +
            theme(plot.title=element_text(hjust=0.5))

grid.arrange(
  plt_base + geom_contour_filled(data=compute_posterior_SB(A_true=0.25, B_true=1),
                                 mapping = aes(x=A, y=B, z=Posterior), show.legend=FALSE) +
             geom_vline(mapping=aes(xintercept = 0.25), colour="firebrick", linetype="dashed", size=1) +
             geom_hline(mapping=aes(yintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Posterior with A/B=0.25 and w=1"),
  plt_base + geom_contour_filled(data=compute_posterior_SB(A_true=0.5, B_true=1),
                                 mapping = aes(x=A, y=B, z=Posterior), show.legend=FALSE) +
             geom_vline(mapping=aes(xintercept = 0.5), colour="firebrick", linetype="dashed", size=1) +
             geom_hline(mapping=aes(yintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Posterior with A/B=0.5 and w=1"),
  plt_base + geom_contour_filled(data=compute_posterior_SB(A_true=1, B_true=1),
                                 mapping = aes(x=A, y=B, z=Posterior), show.legend=FALSE) +
             geom_vline(mapping=aes(xintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             geom_hline(mapping=aes(yintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Posterior with A/B=1 and w=1"),
  plt_base + geom_contour_filled(data=compute_posterior_SB(A_true=2, B_true=1),
                                 mapping = aes(x=A, y=B, z=Posterior), show.legend=FALSE) +
             geom_vline(mapping=aes(xintercept = 2), colour="firebrick", linetype="dashed", size=1) +
             geom_hline(mapping=aes(yintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Posterior with A/B=2 and w=1"),
  plt_base + geom_contour_filled(data=compute_posterior_SB(A_true=4, B_true=1),
                                 mapping = aes(x=A, y=B, z=Posterior), show.legend=FALSE) +
             geom_vline(mapping=aes(xintercept = 4), colour="firebrick", linetype="dashed", size=1) +
             geom_hline(mapping=aes(yintercept = 1), colour="firebrick", linetype="dashed", size=1) +
             labs(title="Posterior with A/B=4 and w=1"),
  nrow=3)
```



