---
title: "Laboratory Session - March 24, 2022"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Paolo Zinesi 2053062

## Exercise 1

Import data

```{r}
data_lakes <- read.csv("scottish_lakes_data.csv", header = TRUE, check.names = FALSE)
data_lakes
```

### Exercise 1.1

Find min & max volume

```{r}
idx_max_vol <- which.max(data_lakes[,"Volume(km3)"])
idx_min_vol <- which.min(data_lakes[,"Volume(km3)"])
lake_max_vol <- data_lakes[idx_max_vol, "Loch"]
lake_min_vol <- data_lakes[idx_min_vol, "Loch"]
max_vol <- data_lakes[idx_max_vol, "Volume(km3)"]
min_vol <- data_lakes[idx_min_vol, "Volume(km3)"]

cat("Lake with largest volume is", lake_max_vol, "with volume =", max_vol, "km^3\n")
cat("Lake with smallest volume is", lake_min_vol, "with volume =", min_vol, "km^3\n")
```

Find min & max area

```{r}
idx_max_area <- which.max(data_lakes[,"Area(km2)"])
idx_min_area <- which.min(data_lakes[,"Area(km2)"])
lake_max_area <- data_lakes[idx_max_area, "Loch"]
lake_min_area <- data_lakes[idx_min_area, "Loch"]
max_area <- data_lakes[idx_max_area, "Area(km2)"]
min_area <- data_lakes[idx_min_area, "Area(km2)"]


cat("Lake with largest area is", lake_max_area, "with area =", max_area, "km^2\n")
cat("Lake with smallest area is", lake_min_area, "with area =", min_area, "km^2\n")
```

### Exercise 1.2

Order dataframe w.r.t. area

```{r}
areasorted_data_lakes <- data_lakes[order(data_lakes[,"Area(km2)"], decreasing = TRUE),]
areasorted_data_lakes
```

```{r}
cat("Lakes with largest areas are", areasorted_data_lakes[1,"Loch"], "and", areasorted_data_lakes[2,"Loch"], ",",
    "with areas", areasorted_data_lakes[1,"Area(km2)"], "and", areasorted_data_lakes[2,"Area(km2)"], "km^2 , respectively")
```

### Exercise 1.3

Sum up the areas occpupied by the lakes to determine the area of Scotland covered by water

```{r}
sum_area <- sum(data_lakes[,"Area(km2)"])
cat("Area of Scotland covered by lakes =", sum_area, "km^2")
```

<br><br>
## Exercise 2

### Exercise 2.1

Import data

```{r}
data_oil <- read.csv("crude-oil-prices.csv", header = TRUE, check.names = FALSE)
data_oil

years <- data_oil[,"Year"]
prices <- data_oil[,"Oil - Crude prices since 1861 (current $)"]
```

### Exercise 2.2

Plot oil price vs years

```{r}
plot(x = years, y = prices, type = 'o',
     main = "Oil price vs years", xlab = "Year", ylab = "Oil price (current $)",
     pch = 20, col = "steelblue", cex = 0.75, lwd = 1)

axis(side = 1, at=seq(1850,2000,25))
```

### Exercise 2.3

Find maximum price in history

```{r}
idx_max_oilprice <- which.max(prices)

cat("The highest price in history is", prices[idx_max_oilprice], "$, it occured in", years[idx_max_oilprice])
```

### Exercise 2.4

Plot price derivative

```{r}
prices_derivative <- (prices[-1] - prices[-(length(prices))])


plot(x = years[-(length(years))], y = prices_derivative, type = 'o',
     main = "Oil price derivative vs years", xlab = "Year", ylab = "Oil price derivative (current $)",
     pch = 20, col = "steelblue", cex = 0.75, lwd = 1)

axis(side = 1, at=seq(1850,2000,25))

```

<br><br>
## Exercise 3

```{r}
library(tibble)
library(readr)
```

### Exercise 3.1

Import data in Tibble format

```{r}
data_coal <- read_csv("coal-production-by-country.csv", show_col_types = FALSE)
data_coal
```

### Exercise 3.2

Count the number of countries available in the file and produce a barplot with the number of entries for each country

```{r}
unique_countries <- unique(data_coal[["Entity"]])
N_countries <- length(unique_countries)
cat("Number of countries available =", N_countries)
```

```{r}
counts_countries <- aggregate(x = list("Counts" = data_coal[["Entity"]]),
                              by =list("Entity" = data_coal[["Entity"]]),
                              FUN = length)
counts_countries
```

```{r, fig.width=5}
barplot(counts_countries[["Counts"]], names.arg = counts_countries[["Entity"]],
        main = "Occurences for each country", xlab = "Country", ylab = "Occurrences", col = "steelblue")
```

### Exercise 3.3

Select only the years ≥ 1970 and compute the total production for each country

```{r}
data_coal_after1970 <- data_coal[data_coal[["Year"]]>=1970,]

productions_after1970 <- aggregate(x = list("Coal production (TWh)" = data_coal_after1970[["Coal production (TWh)"]]),
                                  by =list("Entity" = data_coal_after1970[["Entity"]]),
                                  FUN = sum)

colnames(productions_after1970) <- c("Entity", "Coal production (TWh)")
productions_after1970
```

Print the top 5 countries for production

```{r}
productionsorted_after1970 <- productions_after1970[order(productions_after1970[,"Coal production (TWh)"], decreasing = TRUE),]

productionsorted_after1970[c(1:5),]
```

### Exercise 3.4

Plot production as a function of time for each of the 5 top countries

```{r, fig.height=5, fig.width=5}
par(mfrow=c(3, 2), mar=c(4,4,4,4))

for (i in c(1:5)) {
  ent <- productionsorted_after1970[i,"Entity"]
  mask <- (data_coal_after1970[["Entity"]] == ent)
  
  plot(x = data_coal_after1970[mask, "Year"][[1]],
       y = data_coal_after1970[mask, "Coal production (TWh)"][[1]],
       main = ent, xlab = "Year", ylab = "Coal production (TWh)",
       col = "steelblue", type = 'o', pch = 20, cex = 1.5, lwd = 1.5)
}

```

### Exercise 3.5

Plot cumulative sum of the World’s coal production over the years

```{r, fig.height=3}
yearproduction_after1970 <- aggregate(x = list("Coal production (TWh)" = data_coal_after1970[["Coal production (TWh)"]]),
                                      by =list("Year" = data_coal_after1970[["Year"]]),
                                      FUN = sum)
colnames(yearproduction_after1970) <- c("Year", "Coal production (TWh)")

plot(x = yearproduction_after1970[["Year"]],
     y = cumsum(yearproduction_after1970[["Coal production (TWh)"]]),
     main = "Cumulative production of coal over the years",
     xlab = "Year", ylab = "Coal production (TWh)",
     col = "steelblue", type = 'o', pch = 20, cex = 1.5, lwd = 1.5)
```

<br><br>
## Exercise 4
```{r}
library(dplyr)
```

### Exercise 4.1

Import data
```{r}
url <- "https://raw.githubusercontent.com/owid/covid-19-data/3192ec32ed721ff4efd9081b36c0e6d8406c1114/public/data/vaccinations/vaccinations-by-manufacturer.csv"

data_vaccine <- read_csv(url, show_col_types = FALSE)

data_vaccine
```


### Italy

Filter data for vaccines in Italy

```{r}
# vaccines in sparse order (some values are missing)
italy_vaccines <- filter(data_vaccine, data_vaccine[["location"]] == "Italy")
italy_vaccines
```

```{r}
# useful variables
avail_vaccines <- c("Moderna", "Pfizer/BioNTech", "Johnson&Johnson", "Oxford/AstraZeneca", "Novavax")

datemin <- min(italy_vaccines[["date"]])
datemax <- max(italy_vaccines[["date"]])
dates <- seq(from=datemin, to=datemax, by="days")

vaccmin <- min(italy_vaccines[["total_vaccinations"]])
vaccmax <- max(italy_vaccines[["total_vaccinations"]])
```


Rearrange vaccination data in order to fill al the missing values

```{r}
# full tibble with 0s
df_tmp <- tibble("location"="Italy",
                 "date"=rep(dates, each=length(avail_vaccines)),
                 "vaccine"=rep(avail_vaccines, length(dates)),
                 "total_vaccinations"=0)


# combine original data with 0s and aggregate to remove duplicates
total_italy_vaccines <- tibble(aggregate(total_vaccinations ~ vaccine + date,
                                         data = rbind(df_tmp, italy_vaccines),
                                         FUN = sum))


# loop over data to set 'total_vaccinations' as the previous day if the actual one is zero
for (i in c((length(avail_vaccines)+1):dim(total_italy_vaccines)[1])){
  
  if(total_italy_vaccines[i,"total_vaccinations"] == 0){
    total_italy_vaccines[i,"total_vaccinations"] <- total_italy_vaccines[i-length(avail_vaccines),"total_vaccinations"]
  }
  
}

total_italy_vaccines
```


### Exercise 4.2 (Italy)

Plot number of vaccines as a function of time for different manufacturers (in Italy)

```{r, fig.height=3.5}

# plot loop
for (i in c(1:length(avail_vaccines))) {
  vacc <- avail_vaccines[i]
  mask <- (total_italy_vaccines[["vaccine"]] == vacc)
  
  plot(x = total_italy_vaccines[mask, "date"][[1]],
       y = total_italy_vaccines[mask, "total_vaccinations"][[1]],
       xlim = c(datemin, datemax), ylim = c(vaccmin, vaccmax),
       main = "Vaccine somministrations as a function of time (Italy)",
       xlab = "Date", ylab = "Total vaccinations",
       type = 'p', pch = 20, cex = 0.5, col = i,
       axes = FALSE)
  
  par(new=TRUE)
  
}

# define axis
axis.Date(side=1, at=seq(from=datemin, to=datemax, by="months"), format="%Y-%m")
axis(side=2)

# define legend
legend(x = "topleft", legend = avail_vaccines,
       col = c(1:length(avail_vaccines)), lwd = 3)
```


### Exercise 4.3 (Italy)

Plot total number of vaccines shot per day in Italy

```{r}
dailysum_italy_vaccines <- tibble(aggregate(x = list("total_vaccinations" = total_italy_vaccines[["total_vaccinations"]]),
                                            by = list("date" = total_italy_vaccines[["date"]]),
                                            FUN = sum))
dailysum_italy_vaccines
```

```{r}
# compute day-by-day difference
derivative_italy_vaccines <- tibble("date" = dailysum_italy_vaccines[-1, "date"][[1]],
                                    "daily_vaccinations" = (dailysum_italy_vaccines[-1, "total_vaccinations"][[1]] - dailysum_italy_vaccines[-(dim(dailysum_italy_vaccines)[1]),"total_vaccinations"][[1]]))

derivative_italy_vaccines

```


```{r, fig.height=3.5}
plot(x = derivative_italy_vaccines[["date"]],
     y = derivative_italy_vaccines[["daily_vaccinations"]],
     type = 'o', main = "Daily vaccinations as a function of time (in Italy)",
     xlab = "Date", ylab = "Vaccinations per day",
     pch = 20, col = "steelblue", cex = 0.75, lwd = 1,
     axes = FALSE)

# define axis
axis.Date(side=1, at=dates[c(TRUE, rep(FALSE,20))], format="%Y-%m-%d")
axis(side=2)
```



Let's repeat the same plots for Germany and United States of America:

### Germany

Filter data for vaccines in Germany

```{r}
# vaccines in sparse order (some values are missing)
germany_vaccines <- filter(data_vaccine, data_vaccine[["location"]] == "Germany")
germany_vaccines
```

```{r}
# useful variables
avail_vaccines <- c("Moderna", "Pfizer/BioNTech", "Johnson&Johnson", "Oxford/AstraZeneca", "Novavax")

datemin <- min(germany_vaccines[["date"]])
datemax <- max(germany_vaccines[["date"]])
dates <- seq(from=datemin, to=datemax, by="days")

vaccmin <- min(germany_vaccines[["total_vaccinations"]])
vaccmax <- max(germany_vaccines[["total_vaccinations"]])
```


Rearrange vaccination data in order to fill al the missing values

```{r}
# full tibble with 0s
df_tmp <- tibble("location"="Germany",
                 "date"=rep(dates, each=length(avail_vaccines)),
                 "vaccine"=rep(avail_vaccines, length(dates)),
                 "total_vaccinations"=0)


# combine original data with 0s and aggregate to remove duplicates
total_germany_vaccines <- tibble(aggregate(total_vaccinations ~ vaccine + date,
                                         data = rbind(df_tmp, germany_vaccines),
                                         FUN = sum))


# loop over data to set 'total_vaccinations' as the previous day if the actual one is zero
for (i in c((length(avail_vaccines)+1):dim(total_germany_vaccines)[1])){
  
  if(total_germany_vaccines[i,"total_vaccinations"] == 0){
    total_germany_vaccines[i,"total_vaccinations"] <- total_germany_vaccines[i-length(avail_vaccines),"total_vaccinations"]
  }
  
}

total_germany_vaccines
```


### Exercise 4.2 (Germany)

Plot number of vaccines as a function of time for different manufacturers (in Germany)

```{r, fig.height=3.5}

# plot loop
for (i in c(1:length(avail_vaccines))) {
  vacc <- avail_vaccines[i]
  mask <- (total_germany_vaccines[["vaccine"]] == vacc)
  
  plot(x = total_germany_vaccines[mask, "date"][[1]],
       y = total_germany_vaccines[mask, "total_vaccinations"][[1]],
       xlim = c(datemin, datemax), ylim = c(vaccmin, vaccmax),
       main = "Vaccine somministrations as a function of time (Germany)",
       xlab = "Date", ylab = "Total vaccinations",
       type = 'p', pch = 20, cex = 0.5, col = i,
       axes = FALSE)
  
  par(new=TRUE)
  
}

# define axis
axis.Date(side=1, at=seq(from=datemin, to=datemax, by="months"), format="%Y-%m")
axis(side=2)

# define legend
legend(x = "topleft", legend = avail_vaccines,
       col = c(1:length(avail_vaccines)), lwd = 3)
```


### Exercise 4.3 (Germany)

Plot total number of vaccines shot per day in Germany

```{r}
dailysum_germany_vaccines <- tibble(aggregate(x = list("total_vaccinations" = total_germany_vaccines[["total_vaccinations"]]),
                                            by = list("date" = total_germany_vaccines[["date"]]),
                                            FUN = sum))
dailysum_germany_vaccines
```

```{r}
# compute day-by-day difference
derivative_germany_vaccines <- tibble("date" = dailysum_germany_vaccines[-1, "date"][[1]],
                                    "daily_vaccinations" = (dailysum_germany_vaccines[-1, "total_vaccinations"][[1]] - dailysum_germany_vaccines[-(dim(dailysum_germany_vaccines)[1]),"total_vaccinations"][[1]]))

derivative_germany_vaccines

```


```{r, fig.height=3.5}
plot(x = derivative_germany_vaccines[["date"]],
     y = derivative_germany_vaccines[["daily_vaccinations"]],
     type = 'o', main = "Daily vaccinations as a function of time (in Germany)",
     xlab = "Date", ylab = "Vaccinations per day",
     pch = 20, col = "steelblue", cex = 0.75, lwd = 1,
     axes = FALSE)

# define axis
axis.Date(side=1, at=dates[c(TRUE, rep(FALSE,20))], format="%Y-%m-%d")
axis(side=2)
```

### United States

Filter data for vaccines in the USA

```{r}
# vaccines in sparse order (some values are missing)
usa_vaccines <- filter(data_vaccine, data_vaccine[["location"]] == "United States")
usa_vaccines
```

The data contain many mistakes, here we try to fix the most relevant ones...
```{r}
# FIX BY HAND the biggest mistake in the original data

mask <- (usa_vaccines[["date"]] == '2022-03-16' & usa_vaccines[["vaccine"]] == 'Johnson&Johnson')
usa_vaccines[mask, "total_vaccinations"] <- 0
```

```{r}
# useful variables
avail_vaccines <- c("Moderna", "Pfizer/BioNTech", "Johnson&Johnson")

datemin <- min(usa_vaccines[["date"]])
datemax <- max(usa_vaccines[["date"]])
dates <- seq(from=datemin, to=datemax, by="days")

vaccmin <- min(usa_vaccines[["total_vaccinations"]])
vaccmax <- max(usa_vaccines[["total_vaccinations"]])
```


Rearrange vaccination data in order to fill al the missing values

```{r}
# complete tibble with 0s
df_tmp <- tibble("location"="United States",
                 "date"=rep(dates, each=length(avail_vaccines)),
                 "vaccine"=rep(avail_vaccines, length(dates)),
                 "total_vaccinations"=0)


# combine original data with 0s and aggregate to remove duplicates
total_usa_vaccines <- tibble(aggregate(total_vaccinations ~ vaccine + date,
                                         data = rbind(df_tmp, usa_vaccines),
                                         FUN = sum))


# loop over data to set 'total_vaccinations' as the previous day if the actual one is zero
for (i in c((length(avail_vaccines)+1):dim(total_usa_vaccines)[1])){
  
  if(total_usa_vaccines[i,"total_vaccinations"] == 0){
    total_usa_vaccines[i,"total_vaccinations"] <- total_usa_vaccines[i-length(avail_vaccines),"total_vaccinations"]
  }
  
}


total_usa_vaccines
```


### Exercise 4.2 (United States)

Plot number of vaccines as a function of time for different manufacturers (in the United States)

```{r, fig.height=3.5}

# plot loop
for (i in c(1:length(avail_vaccines))) {
  vacc <- avail_vaccines[i]
  mask <- (total_usa_vaccines[["vaccine"]] == vacc)
  
  plot(x = total_usa_vaccines[mask, "date"][[1]],
       y = total_usa_vaccines[mask, "total_vaccinations"][[1]],
       xlim = c(datemin, datemax), ylim = c(vaccmin, vaccmax),
       main = "Vaccine somministrations as a function of time (United States)",
       xlab = "Date", ylab = "Total vaccinations",
       type = 'p', pch = 20, cex = 0.5, col = i,
       axes = FALSE)
  
  par(new=TRUE)
  
}

# define axis
axis.Date(side=1, at=seq(from=datemin, to=datemax, by="months"), format="%Y-%m")
axis(side=2)

# define legend
legend(x = "topleft", legend = avail_vaccines,
       col = c(1:length(avail_vaccines)), lwd = 3)
```


### Exercise 4.3 (United States)

Plot total number of vaccines shot per day in United States

```{r}
dailysum_usa_vaccines <- tibble(aggregate(x = list("total_vaccinations" = total_usa_vaccines[["total_vaccinations"]]),
                                          by = list("date" = total_usa_vaccines[["date"]]),
                                          FUN = sum))
dailysum_usa_vaccines
```

```{r}
# compute day-by-day difference
derivative_usa_vaccines <- tibble("date" = dailysum_usa_vaccines[-1, "date"][[1]],
                                    "daily_vaccinations" = (dailysum_usa_vaccines[-1, "total_vaccinations"][[1]] - dailysum_usa_vaccines[-(dim(dailysum_usa_vaccines)[1]),"total_vaccinations"][[1]]))

derivative_usa_vaccines

```


```{r, fig.height=3.5}
plot(x = derivative_usa_vaccines[["date"]],
     y = derivative_usa_vaccines[["daily_vaccinations"]],
     type = 'o', main = "Daily vaccinations as a function of time (in the United States)",
     xlab = "Date", ylab = "Vaccinations per day",
     pch = 20, col = "steelblue", cex = 0.75, lwd = 1,
     axes = FALSE)

# define axis
axis.Date(side=1, at=dates[c(TRUE, rep(FALSE,20))], format="%Y-%m-%d")
axis(side=2)
```


### Exercise 4.4

Country-by-country data on global COVID-19 vaccinations
```{r}
url <- "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv"

data_vaccine_countries <- read_csv(url, show_col_types = FALSE)

data_vaccine_countries
```

Filter data for vaccines in Europe
```{r}
europe_vaccines <- filter(data_vaccine_countries, data_vaccine_countries[["iso_code"]] == "OWID_EUR")
europe_vaccines
```
Plot the number of daily vaccinations per million as a function of date
```{r, fig.height=3.5}
datemin <- min(europe_vaccines[["date"]])
datemax <- max(europe_vaccines[["date"]])
dates <- seq(from=datemin, to=datemax, by="days")


plot(x = europe_vaccines[["date"]],
     y = europe_vaccines[["daily_vaccinations_per_million"]],
     type = 'o', main = "Daily vaccinations as a function of time (in Europe)",
     xlab = "Date", ylab = "Vaccinations per million people per day",
     pch = 20, col = "steelblue", cex = 0.75, lwd = 1,
     axes = FALSE)

# define axis
axis.Date(side=1, at=dates[c(TRUE, rep(FALSE,30))], format="%Y-%m-%d")
axis(side=2)
```


### Exercise 4.5

Produce cool plots
```{r}
# select all european countries
url <- "https://raw.githubusercontent.com/ajturner/acetate/master/places/Countries-Europe.csv"
european_countries <- read_csv(url, show_col_types = FALSE)
european_countries <- european_countries[, c("name", "ISO alpha 3")]
colnames(european_countries) <- c("Country", "ISO_code")

european_countries
```
```{r}

european_countries_vaccinations <- filter(data_vaccine_countries,
                                          ifelse(data_vaccine_countries[["iso_code"]] %in% european_countries[["ISO_code"]], TRUE, FALSE))


european_countries_vaccinations[european_countries_vaccinations[["date"]] %in% seq(from=as.Date("2022-03-20"),to=as.Date("2022-03-28"), by="days"),]

```

Here we plot the comparison between single european countries and the average of all of them
```{r, fig.height=3.5}
# useful parameters
datemin <- min(european_countries_vaccinations[["date"]])
datemax <- max(european_countries_vaccinations[["date"]])
dates <- seq(from=datemin, to=datemax, by="days")
vaccmin <- min(european_countries_vaccinations[["daily_vaccinations_per_million"]], na.rm = TRUE)
vaccmax <- max(european_countries_vaccinations[["daily_vaccinations_per_million"]], na.rm = TRUE)

# plot vaccinations of all countries in Europe
plot(x = european_countries_vaccinations[["date"]],
     y = european_countries_vaccinations[["daily_vaccinations_per_million"]],
     xlim = c(datemin, datemax), ylim = c(vaccmin, vaccmax),
     xlab = "Date", ylab = "Vaccinations per million people per day",
     type = 'o', main = "Daily vaccinations as a function of time (single european countries vs european average)",
     pch = 20, cex = 0.75, lwd = 1,
     axes = FALSE)

par(new=TRUE)

# plot (mean) vaccinations of Europe
plot(x = europe_vaccines[["date"]],
     y = europe_vaccines[["daily_vaccinations_per_million"]],
     xlim = c(datemin, datemax), ylim = c(vaccmin, vaccmax),
     xlab = "", ylab = "",
     type = 'o',
     pch = 20, col = "red", cex = 0.75, lwd = 1,
     axes = FALSE)

# define axis
axis.Date(side=1, at=dates[c(TRUE, rep(FALSE,30))], format="%Y-%m-%d")
axis(side=2)

# define legend
legend(x = "topright", legend = c("European countries","Europe (average)"),
       col = c(1,2), lwd = 3)
```

Plot time progression of number of vaccinated people for each vaccination level
```{r, fig.height=3.5}
# useful parameters
datemin <- min(europe_vaccines[["date"]])
datemax <- max(europe_vaccines[["date"]])
dates <- seq(from=datemin, to=datemax, by="days")
vaccmin <- min(europe_vaccines[["people_vaccinated"]], na.rm = TRUE)
vaccmax <- max(europe_vaccines[["people_vaccinated"]], na.rm = TRUE)

avail_feat <- c("people_vaccinated", "people_fully_vaccinated", "total_boosters")

# plot loop
for (i in c(1:length(avail_feat))) {
  feat <- avail_feat[i]
  
  plot(x = europe_vaccines[, "date"][[1]],
       y = europe_vaccines[, feat][[1]],
       xlim = c(datemin, datemax), ylim = c(vaccmin, vaccmax),
       main = "Vaccinated people for each vaccination level (in Europe)",
       xlab = "Date", ylab = "Vaccinations",
       type = 'p', pch = 20, cex = 0.5, col = i+1,
       axes = FALSE)
  
  par(new=TRUE)
  
}

# define axis
axis.Date(side=1, at=dates[c(TRUE, rep(FALSE,30))], format="%Y-%m-%d")
axis(side=2)

# define legend
legend(x = "topleft", legend = c("Vaccinated", "Fully vaccinated", "Booster-vaccinated"),
       col = c(2:(length(avail_feat)+1)), lwd = 3)
```


Here we plot the difference between 1st-2nd dose and 2nd-booster dose number of vaccinations
```{r, fig.height=3, fig.width=3}
par(mar=c(5, 4, 4, 4))

# correlation between first-second dose and second-booster dose choices
european_countries_actualvaccinations <- tibble(aggregate(cbind(people_vaccinated, people_fully_vaccinated, total_boosters) ~ iso_code,
                                         data = european_countries_vaccinations,
                                         FUN = max))

# useful variables
xmax <- max(european_countries_actualvaccinations[, avail_feat[2]][[1]])
ymax <- max(european_countries_actualvaccinations[, avail_feat[1]][[1]],
            european_countries_actualvaccinations[, avail_feat[3]][[1]])


# fist-vs-second dose
plot(x = european_countries_actualvaccinations[, avail_feat[2]][[1]],
     y = european_countries_actualvaccinations[, avail_feat[1]][[1]],
     xlim = c(0, xmax), ylim = c(0, ymax),
     xlab = "Fully vaccinated", ylab = "",
     main = "Comparison of 1st-2nd and 2nd-booster vaccinations",
     type = 'p', pch = 20, cex = 1, lwd = 1,
     col = "steelblue",
     axes = FALSE)

par(new=TRUE)

# second-vs-booster dose
plot(x = european_countries_actualvaccinations[, avail_feat[2]][[1]],
     y = european_countries_actualvaccinations[, avail_feat[3]][[1]],
     xlim = c(0, xmax), ylim = c(0, ymax),
     xlab = "Fully vaccinated", ylab = "",
     type = 'p', pch = 20, cex = 1, lwd = 1,
     col = "firebrick",
     axes = FALSE)

# define axis and axis labels
axis(1)
axis(2, col.axis="steelblue")
axis(4, col.axis="firebrick")
mtext("Vaccinated", side=2, line=3, cex.lab=1,las=0, col="steelblue")
mtext("Booster-vaccinated", side=4, line=3, cex.lab=1,las=0, col="firebrick")
```
